---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# uncertain

<!-- badges: start -->
<!-- badges: end -->

`uncertain` is a package for performing uncertainty of measurement calculations. 
These calculations are used in the validation of genetic laboratory tests to 
communicate the variability within an assay.

However, the specific calculations required are sometimes unclear, and there is
a lack of worked examples relating to genomic testing. 
`uncertain` aims to address this by providing functions for key calculations, and
example datasets which reflect the data structures encountered in 
genomic laboratories.

## Calculating uncertainty

`uncertain` is based on the instructions for pooled standard 
deviation analysis described in the publication "Selected Laboratory and Measurement 
Practices and Procedures to Support Basic Mass Calibrations" from the National
Institute of Standards and Technology 
([NISTIR 6969](https://nvlpubs.nist.gov/nistpubs/ir/2019/NIST.IR.6969-2019.pdf)).
The functions are tested using example calculations from the 
[NIST Engineering Statistics Handbook](https://www.itl.nist.gov/div898/handbook/mpc/section4/mpc441.htm).

`uncertain` includes example datasets for copy number variant and single
nucleotide variant repeatability data.

```{r}
#| warning: FALSE

library(uncertain)
library(ggplot2)

ggplot(data_cnv, aes(x = gene, y = dosage)) +
  geom_jitter(shape = 21, width = 0.1) +
  theme_bw() +
  facet_wrap(~sample) +
  labs(title = "CNV testing repeat data",
       x = "Gene", y = "Dosage") +
  scale_y_continuous(limits = c(0, 10),
                     breaks = seq(0, 10, by = 2))

```

The key function is `group_stats` which calculates statistical variation within
the dataset based on grouping variables (i.e. sample, gene) provided by the
user.

```{r}

cnv_variation <- group_stats(data_cnv,
            measurement_variable = dosage,
            sample, gene)

knitr::kable(cnv_variation[[1]])

```

`group_stats` also calculates the pooled standard deviation for the whole
dataset.

```{r}

cnv_variation[[2]]

```

## Installation

You can install the development version of uncertain from [GitHub](https://github.com/) with:

``` r
# install.packages("pak")
pak::pak("joe-m-shaw/uncertain")
```
