---
title: "Quantifying assay variation"
vignette: >
  %\VignetteIndexEntry{Quantifying assay variation}
  %\VignetteEngine{quarto::html}
  %\VignetteEncoding{UTF-8}
knitr:
  opts_chunk:
    collapse: true
    comment: '#>'
---

```{r}
#| label: setup
#| include: FALSE

library(uncertain)
library(ggplot2)

```

`uncertain` contains fake datasets to help with calculating assay variation.
For example, data_cnv includes made-up data for an assay which measures the 
dosage of 3 genes (*BRCA1*, *CDKN2A* and *EGFR*) for 3 samples, each of which 
has been tested 6 times.

```{r}
#| include: TRUE

head(data_cnv)

```

```{r}
#| echo: FALSE

ggplot(data_cnv, aes(x = gene, y = dosage)) +
  geom_jitter(shape = 21, width = 0.1) +
  theme_bw() +
  facet_wrap(~sample) +
  labs(title = "CNV testing repeat data",
       x = "Gene", y = "Dosage") +
  scale_y_continuous(limits = c(0, 10),
                     breaks = seq(0, 10, by = 2))

```

A key aspect of assay development is determining the variability within the 
assay. 
`uncertain` contains the `group_stats` function which calculates the statistical
variation within a dataset based on groupings supplied by the user.

```{r}

knitr::kable(group_stats(data_cnv,
            measurement_variable = dosage,
            sample, gene)[[1]])

```

`group_stats` also calculates the pooled standard deviation across all the sample
groupings, which quantifies the overall variation within the assay itself.

```{r}

group_stats(data_cnv,
            measurement_variable = dosage,
            sample, gene)[[2]]

```

The grouping step is very important. If the data are not grouped at the gene
level, then biological variation is misinterpreted as assay variation.
For example, the *CDKN2A* deletion within sample 2 and the *EGFR* amplification 
within sample 3 leads to higher
standard deviations calculated for these samples compared to sample 1, and the 
pooled standard deviation is higher.

```{r}

knitr::kable(group_stats(data_cnv,
            measurement_variable = dosage,
            sample)[[1]])

group_stats(data_cnv,
            measurement_variable = dosage,
            sample)[[2]]
```





